---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# Introduction to ampir

<!-- badges: start -->
[![Build Status](https://travis-ci.com/Legana/ampir.svg?token=fesxqj9vWJzeRTtyzLHt&branch=master)](https://travis-ci.com/Legana/ampir)
[![Travis build status](https://travis-ci.org/Legana/ampir.svg?branch=master)](https://travis-ci.org/Legana/ampir)
<!-- badges: end -->

The **ampir** (short for **a**nti**m**icrobial **p**eptide prediction **i**n **r** ) package was designed to be an open and user-friendly method to predict AMPs (antimicrobial peptides) from any given size protein dataset. **ampir** uses four sequential functions to create a four step approach to predict AMPs.

## Installation

You can install the development version of ampir from [GitHub](https://github.com/) with:

Note: this does not yet work as the repository is private.
```r
# install.packages("devtools")
devtools::install_github("Legana/ampir")
```

```{r, echo = FALSE, eval=FALSE}
#devtools::install_github("Legana/ampir",auth_token = "07d66396f0c3f7a5ac675b8e7312b9eda55d1d8b")
```

## Brief background

**ampir** uses a *supervised statistical machine learning* approach to predict AMPs. Basically this involves making a statistical model based on *input* data to predict *output* data [James, Witten, Hastie & Tibshirani 2013](http://www-bcf.usc.edu/~gareth/ISL/). The input data are also known as *features* which are used to describe the data. To predict AMPs, physicochemical and compositional properties of protein sequences are used as features [Osorio, Rondón-Villarreal & Torres 2015](https://journal.r-project.org/archive/2015/RJ-2015-001/RJ-2015-001.pdf). Therefore, within this package, it is important to ***follow an order of functions***.

### Order of functions to follow

1. `read_faa()` to read FASTA amino acid files.
2. `remove_nonstandard_aa()` to remove non standard amino acid sequences.
3. `calculate_features()` to calculate data descriptors used by the predictive model.
4. `predict_AMP_prob()` to predict the AMP probability of a protein.

## Example workflow

```{r setup, warning=FALSE, message=FALSE}
library(ampir)
```

### Step 1: Read FASTA amino acid files with `read_faa()` 

`read_faa()` reads FASTA amino acid files as a dataframe.

```{r}
my_protein <- read_faa(system.file("extdata/bat_protein.fasta", package = "ampir"))
```

```{r, echo=FALSE}
#sapply to shorten words in df and use data.frame(as.list) to tidy up kable output
#knitr::kable(as.data.frame(as.list(sapply(my_protein, strtrim, 35))), caption = "My protein")

knitr::kable(my_protein, caption = "My protein")
```

### Step 2: Remove non standard amino acids with `remove_nonstandard_aa()`

`remove_nonstandard_aa()` is used to remove sequences that contain anything other than the 20 standard amino acids. These sequences are removed to circumvent potential complications with calculations.


```{r, echo=FALSE}
nonstandard_example_df <- readRDS(system.file("extdata/non_standard_df.rds", package = "ampir"))

#knitr::kable(sapply(nonstandard_example_df, strtrim, 35), caption = "Example dataframe with a nonsense protein")

knitr::kable(nonstandard_example_df, caption = "Example dataframe with a nonsense protein")
```

The table above shows a dataframe with protein sequences in it. The second row contains a made up sequence to serve as an example. This made up sequence will be removed with `remove_nonstandard_aa()`.

```{r}
my_clean_protein <- remove_nonstandard_aa(df = nonstandard_example_df)
```

```{r, echo=FALSE}
#knitr::kable(as.data.frame(as.list(sapply(my_clean_protein, strtrim, 35))), caption = "My clean protein")

knitr::kable(my_clean_protein, caption = "My clean protein")
```


### Step 3: Calculate features with `calculate_features()`

`calculate_features()` calculates a range of physicochemical properties that are used by the predictive model within `predict_AMP_prob()` to make its predictions (step 4).

```{r}
my_protein_features <- calculate_features(df = my_clean_protein)
```

```{r, echo=FALSE}
#knitr::kable(my_protein_features[1:6], digits = 3, caption = "My protein features ( first six columns )")

knitr::kable(my_protein_features, digits = 3, caption = "My protein features")

```

### Step 4: Predict antimicrobial peptide probability with `predict_AMP_prob()` 

`predict_AMP_prob()` uses the output from `calculate_features()` as a parameter to predict the probability of a protein to be an antimicrobial peptide.

```{r}
my_prediction <- predict_AMP_prob(df = my_protein_features)
```

```{r, echo=FALSE}
knitr::kable(my_prediction, digits = 3, caption = "My protein prediction")
```

#### Optional steps: Extract AMP sequences and save as FASTA format file

`extract_amps()` uses the output from `read_faa()` and `predict_AMP_prob()` as parameters to create a new dataframe which contains the sequence name and sequence of the identified antimicrobial peptides at a set probability threshold of >=0.50. The default threshold of >=0.50 can be changed with the "prob" parameter.

```{r}
my_predicted_amps <- extract_amps(df_w_seq = my_protein, df_w_prob = my_prediction, prob = 0.55)
```

```{r, echo=FALSE}
knitr::kable(my_predicted_amps, caption = "My predicted AMPs")
```

`df_to_faa()` writes a dataframe containing the sequence and corresponding sequence name to a FASTA file.

```{r,eval=FALSE}
df_to_faa(my_predicted_amps, "my_predicted_amps.fasta")
```


