---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Introduction to ampir

The **ampir** (short for **a**nti**m**icrobial **p**eptide prediction **i**n **r** ) package was designed to be a fast and user-friendly method to predict AMPs (antimicrobial peptides) from large protein dataset. **ampir** uses a *supervised statistical machine learning* approach to predict AMPs. It incorporates a support vector machine classification model that has been trained on publicly available antimicrobial peptide data. 

<!-- badges: start -->
#[![Travis build status](https://travis-ci.org/Legana/ampir.svg?branch=master)](https://travis-ci.org/Legana/ampir)
<!-- badges: end -->

## Installation

You can install the development version of ampir from [GitHub](https://github.com/) with:

```r
# install.packages("devtools")
devtools::install_github("Legana/ampir")
```


```{r setup, warning=FALSE, message=FALSE}
library(ampir)
```

Standard input to **ampir** is a `data.frame` with sequence names in the first column and protein sequences in the second column.  A convenience function `read_faa()` is provided to create input data by reading a FASTA formatted file. 

```{r, warning=FALSE, message=FALSE}
my_protein_df <- read_faa(system.file("extdata/bat_protein.fasta", package = "ampir"))
```

```{r, echo=FALSE}
display_df <- my_protein_df
display_df$seq_aa <- paste(substring(display_df$seq_aa,1,45),"...",sep="")
knitr::kable(display_df)
```


Calculate the probability that each protein is an antimicrobial peptide with `predict_amps()` 

```{r}
my_prediction <- predict_amps(my_protein_df)
```

```{r, echo=FALSE}
my_prediction$seq_aa <- paste(substring(my_prediction$seq_aa,1,45),"...",sep="")
knitr::kable(my_prediction, digits = 3)
```
