---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# Introduction to ampir

The **ampir** (short for **a**nti**m**icrobial **p**eptide prediction **i**n **r** ) package was designed to be a fast and user-friendly method to predict AMPs (antimicrobial peptides) from large protein dataset. **ampir** uses a *supervised statistical machine learning* approach to predict antimicrobial peptides (AMPs).  It comes with a statistical model that has been trained using publicly available sequence data for known antimicrobial peptides. 

**ampir** uses protein sequences as input and produces a table with the sequence names and the probability of that sequence to be an AMP as output. 


<!-- badges: start -->
[![Build Status](https://travis-ci.com/Legana/ampir.svg?token=fesxqj9vWJzeRTtyzLHt&branch=master)](https://travis-ci.com/Legana/ampir)
<!-- badges: end -->

```{r, echo=FALSE, eval=FALSE}
#[![Travis build status](https://travis-ci.org/Legana/ampir.svg?branch=master)](https://travis-ci.org/Legana/ampir)
```

## Installation

You can install the development version of ampir from [GitHub](https://github.com/) with:

```r
# install.packages("devtools")
devtools::install_github("Legana/ampir")
```


```{r setup, warning=FALSE, message=FALSE}
library(ampir)
```


```

```{r, echo=FALSE}
```



```{r}
```

```{r, echo=FALSE}
```

