---
title: "Introduction to ampir"
author: "Legana Fingerhut"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to ampir}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Background

The **ampir** (short for **a**nti**m**icrobial **p**eptide prediction **i**n **r** ) package was designed to be a fast and user-friendly method to predict AMPs (antimicrobial peptides) from any given size protein dataset. **ampir** uses a *supervised statistical machine learning* approach to predict antimicrobial peptides (AMPs).  It comes with a model that has been trained on publicly available antimicrobial peptide data. 

## Usage

```{r setup, warning=FALSE, message=FALSE}
library(ampir)
```

Read protein sequences from a FASTA formatted file with the function `read_faa()`. 

```{r, warning=FALSE, message=FALSE}
my_protein <- read_faa(system.file("extdata/bat_protein.fasta", package = "ampir"))
```

Calculate the probability that each protein is an antimicrobial peptide with `predict_amps()` 

```{r}
my_prediction <- predict_amps(my_protein)
```

```{r, echo=FALSE}
knitr::kable(my_prediction, digits = 3, caption = "My prediction")
```


